---
title: "AI-net-stats"
output: html_document
---
```{r}
library(jsonlite)
library(RJSONIO)
library(ggplot2)
library(scales)
library(dplyr)
library(tidyverse)
library(forcats)
library(ggh4x)
library(effsize)
library(glmnet)
library(pROC)

set.seed(0)
```

Load metrics
```{r}
ai_metrics <- fromJSON("./inputs/ai_metrics_single_period.json")
human_metrics <- fromJSON("./inputs/human_metrics_single_period.json")


Claude_Code_metrics <- fromJSON("./inputs/Claude_Code_metrics_single_period.json")
Copilot_metrics <- fromJSON("./inputs/Copilot_metrics_single_period.json")
Cursor_metrics <- fromJSON("./inputs/Cursor_metrics_single_period.json")
Devin_metrics <- fromJSON("./inputs/Devin_metrics_single_period.json")
OpenAI_Codex_metrics <- fromJSON("./inputs/OpenAI_Codex_metrics_single_period.json")
```

1. [RQ1] Plot period metrics
  Extract period data
```{r}
ai_period <- data.frame(ai_metrics$per_period[[1]][3:30])
ai_period$count <- length(ai_metrics$per_issue)
ai_period$type <- "AI"

human_period <- data.frame(human_metrics$per_period[[1]][3:30])
human_period$count <- length(human_metrics$per_issue)
human_period$type <- "Human"

Claude_Code_period <- data.frame(Claude_Code_metrics$per_period[[1]][3:30])
Claude_Code_period$count <- length(Claude_Code_metrics$per_issue)
Claude_Code_period$type <- "Claude Code"

Copilot_period <- data.frame(Copilot_metrics$per_period[[1]][3:30])
Copilot_period$count <- length(Copilot_metrics$per_issue)
Copilot_period$type <- "Copilot"

Cursor_period <- data.frame(Cursor_metrics$per_period[[1]][3:30])
Cursor_period$count <- length(Cursor_metrics$per_issue)
Cursor_period$type <- "Cursor"

Devin_period <- data.frame(Devin_metrics$per_period[[1]][3:30])
Devin_period$count <- length(Devin_metrics$per_issue)
Devin_period$type <- "Devin"

OpenAI_Codex_period <- data.frame(OpenAI_Codex_metrics$per_period[[1]][3:30])
OpenAI_Codex_period$count <- length(OpenAI_Codex_metrics$per_issue)
OpenAI_Codex_period$type <- "OpenAI Codex"

all_period <- bind_rows(ai_period, human_period, Claude_Code_period, Copilot_period,
                        Cursor_period, Devin_period, OpenAI_Codex_period)
```

  Create radio plot data
```{r}
all_period_ratio <- all_period %>% mutate(across(c("edges", "vertices"), ~ .x / count))

all_period_ratio <- all_period_ratio %>% mutate(across(-type, ~ .x / .x[type == "Human"])) %>% pivot_longer(-type)

all_period_ratio <- all_period_ratio %>%
  mutate(category = case_when(
    str_detect(name, "sum") ~ "SUMS",
    str_detect(name, "constraint") ~ "Constraint Avg",
    str_detect(name, "betweenness") ~ "Betweenness Avg",
    str_detect(name, "closeness") ~ "Closeness Avg",
    str_detect(name, "effective_size") ~ "Effective Size Avg",
    str_detect(name, "efficiency") ~ "Efficiency Avg",
    str_detect(name, "hierarchy") ~ "Hierarchy Avg",
    str_detect(name, "edges") ~ "Average Edges",
    str_detect(name, "vertices") ~ "Average Vertices",
    TRUE ~ str_to_title(name)
  )) %>% filter(str_detect(name, "avg") | name %in% c("edges", "vertices", "density", "diameter"))

all_period_ratio <- all_period_ratio %>% group_by(category) %>% arrange(max(value))

all_period_ratio$category <- reorder(all_period_ratio$category, all_period_ratio$value, FUN = mean)
```

```{r}
ratio_plot <- ggplot(all_period_ratio) +
  geom_vline(xintercept = 1, color = "red", alpha = 0.5) +
  geom_point(aes(x = value, y = category, color=type), size = 3) +
  scale_x_log10(labels = label_number()) +
  labs(x = "Ratio of Human's Value",
       y = "Metric",
       color = "Agent") +
  theme_classic()

show(ratio_plot)
ggsave(filename="./plots/metrics_ratio_plot.png", plot = ratio_plot)
```

  Unused plots
```{r}
all_period_reconfig <- all_period %>% pivot_longer(-type)

all_period_reconfig <- all_period_reconfig %>% group_by(type) %>% mutate(value = if_else(name %in% c("edges", "vertices", "count"), value/value[name=="count"], value))

all_period_reconfig <- all_period_reconfig %>%
  mutate(category = case_when(
    str_detect(name, "sum") ~ "SUMS",
    str_detect(name, "constraint") ~ "Constraint",
    str_detect(name, "betweenness") ~ "Betweenness",
    str_detect(name, "closeness") ~ "Closeness",
    str_detect(name, "effective_size") ~ "Effective Size",
    str_detect(name, "efficiency") ~ "Efficiency",
    str_detect(name, "hierarchy") ~ "Hierarchy",
    str_detect(name, "edges") ~ "Average Edges",
    str_detect(name, "vertices") ~ "Average Vertices",
    TRUE ~ str_to_title(name)
  )) %>% filter(str_detect(name, "avg") | name %in% c("edges", "vertices", "density", "diameter"))

all_period_reconfig$category <- factor(all_period_reconfig$category, levels=unique(all_period_reconfig$category))
```

```{r}
scales <- list(
  all_period_reconfig$name == "edges" ~ scale_y_continuous(limits = c(0, 10)),
  all_period_reconfig$name == "diameter" ~ scale_y_continuous(limits = c(0, 10)),
  TRUE ~ scale_y_continuous(limits = c(0,1))
)

metrics_plot <- ggplot(all_period_reconfig %>% filter(type %in% c("AI", "Human")), aes(x=name, y = value, fill=type)) +
  geom_col(position="dodge") + 
  facet_wrap(~category, scales = "free") +
  facetted_pos_scales(y = scales) +
  scale_fill_brewer(palette="Pastel2") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

show(metrics_plot)
ggsave(filename="./plots/AIxHuman_metrics.png", plot = metrics_plot, height = 16, width = 12)
```
```{r}
metrics_plot <- ggplot(all_period_reconfig %>% filter(!type %in% c("AI", "Human")), aes(x=name, y = value, fill=type)) +
  geom_col(position="dodge") + 
  facet_wrap(~category, scales = "free") +
  facetted_pos_scales(y = scales) +
  scale_fill_brewer(palette="Pastel1") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

show(metrics_plot)
ggsave(filename="./plots/agents_metrics.png", plot = metrics_plot, height = 16, width = 12)
```

2. [RQ1] Perform Mann-Whitney U test and Cliff Delta
  Unpack issues data
```{r}
ai_issues <- as.data.frame(do.call(rbind, ai_metrics$per_issue))
ai_issues <- cbind(ai_issues, do.call(rbind, ai_metrics$per_period[[1]]$per_period_issue))
ai_issues$group <- "AI"

human_issues <- as.data.frame(do.call(rbind, human_metrics$per_issue))
human_issues <- cbind(human_issues, do.call(rbind, human_metrics$per_period[[1]]$per_period_issue))
human_issues$group <- "Human"

Claude_Code_issues <- as.data.frame(do.call(rbind, Claude_Code_metrics$per_issue))
Claude_Code_issues <- cbind(Claude_Code_issues, do.call(rbind, Claude_Code_metrics$per_period[[1]]$per_period_issue))
Claude_Code_issues$group <- "Claude Code"

Copilot_issues <- as.data.frame(do.call(rbind, Copilot_metrics$per_issue))
Copilot_issues <- cbind(Copilot_issues, do.call(rbind, Copilot_metrics$per_period[[1]]$per_period_issue))
Copilot_issues$group <- "Copilot"

Cursor_issues <- as.data.frame(do.call(rbind, Cursor_metrics$per_issue))
Cursor_issues <- cbind(Cursor_issues, do.call(rbind, Cursor_metrics$per_period[[1]]$per_period_issue))
Cursor_issues$group <- "Cursor"

Devin_issues <- as.data.frame(do.call(rbind, Devin_metrics$per_issue))
Devin_issues <- cbind(Devin_issues, do.call(rbind, Devin_metrics$per_period[[1]]$per_period_issue))
Devin_issues$group <- "Devin"

OpenAI_Codex_issues <- as.data.frame(do.call(rbind, OpenAI_Codex_metrics$per_issue))
OpenAI_Codex_issues <- cbind(OpenAI_Codex_issues, do.call(rbind, OpenAI_Codex_metrics$per_period[[1]]$per_period_issue))
OpenAI_Codex_issues$group <- "OpenAI Codex"

all_issues <- rbind(ai_issues, human_issues, Claude_Code_issues, Copilot_issues,
                    Cursor_issues, Devin_issues, OpenAI_Codex_issues)

all_issues$density[all_issues$density == "NULL"] <- NA
all_issues$merged_at[all_issues$merged == "NULL"] <- NA

all_issues$group <- as.factor(all_issues$group)
all_issues$merged <- ifelse(is.na(all_issues$merged_at), 0, 1)

all_issues <- all_issues %>% select(-c("participants", "agent", "state", "html_url", "merged_at"))

all_issues <- all_issues %>% mutate(across(-group, as.numeric))
```

  Run MW on each group pair for each metric
```{r}
MWC_U_tests <- data.frame()
pairs_df <- data.frame(combn(unique(as.character(all_issues$group)), 2))

for(col_name in colnames(pairs_df)){
  #print(col_name)
  pair <- pairs_df[[col_name]]
  #print(pair)
  MWC_pair <- lapply(select((all_issues %>% filter(group %in% pair)), -group, -merged),
                     function(x){
                       wilcox.test(
                         x ~ (all_issues %>% filter(group %in% pair))$group,
                         alternative="two.sided")
                     }
  )
  
  MWC_pair <- as.data.frame(do.call(rbind, MWC_pair))
  MWC_U_tests <- rbind(MWC_U_tests, format(round(as.numeric(MWC_pair$p.value), 5), nsmall = 5, scientific = FALSE))
  rownames(MWC_U_tests)[nrow(MWC_U_tests)] <- sprintf("%s x %s", pair[1], pair[2])
  colnames(MWC_U_tests) <- rownames(MWC_pair)
  MWC_U_tests[] <- lapply(MWC_U_tests, as.numeric)
}
```

  Run cliff delta on each group pair for each metric
```{r}
cliff_tests <- data.frame()

for(col_name in colnames(pairs_df)){
  #print(col_name)
  pair <- pairs_df[[col_name]]
  #print(pair)
  cliff_pair <- lapply(select((all_issues %>% filter(group %in% pair)), -group, -merged),
                       function(x){
                         cliff.delta(
                           x ~ relevel(droplevels((all_issues %>% 
                                                     filter(group %in% pair))$group),
                                       pair[1])
                         )
                       }
  )
  
  cliff_pair <- as.data.frame(do.call(rbind, cliff_pair))
  cliff_tests <- rbind(cliff_tests, format(round(as.numeric(cliff_pair$estimate), 2), nsmall = 2, scientific = FALSE))
  rownames(cliff_tests)[nrow(cliff_tests)] <- sprintf("%s x %s", pair[1], pair[2])
  colnames(cliff_tests) <- rownames(cliff_pair)
  cliff_tests[] <- lapply(cliff_tests, as.numeric)
}
```

  Combine into a single table
```{r}
cliff_table <- mapply(paste, 
                      cliff_tests,
                      cliff_tests %>% mutate(across(everything(), ~ case_when(
                        abs(.x) < 0.147 ~ "negligible",
                        abs(.x) < 0.33 ~ "small",
                        abs(.x) < 0.474 ~ "medium",
                        TRUE ~ "large"))),
                      MWC_U_tests %>% mutate(across(everything(), ~ case_when(
                        .x < 0.001 ~ "***",
                        .x < 0.01 ~ "**",
                        .x < 0.05 ~ "*",
                        TRUE ~ ""))),
                      sep = " "
)
rownames(cliff_table) <- rownames(cliff_tests)
```

```{r}
#write.csv(MWC_U_tests, "Mann-Whitney_U_tests.csv")
#write.csv(cliff_tests, "cliff_tests.csv")
write.csv(cliff_table, "cliff_table.csv")
```

3. [RQ2] Run regularization over issue metrics

  Check baseline truth
```{r}
  sprintf("Human: %f", sum((all_issues %>% filter(group == "Human"))$merged) / nrow(all_issues %>% filter(group == "Human")))
sprintf("AI: %f", sum((all_issues %>% filter(group == "AI"))$merged) / nrow(all_issues %>% filter(group == "AI")))
```
  Check human alone
```{r}
set.seed(0)
norm_issues <- all_issues %>% filter(group %in% c("Human")) %>%mutate(across(where(is.numeric) & !merged, ~ as.vector(scale(.))))
norm_issues$group <- droplevels(norm_issues$group)

x <- model.matrix(merged ~ ., data = norm_issues %>% select(-vertices, -group))[,-1]
y <- (norm_issues)$merged
cv <- cv.glmnet(x, y, alpha = 0, family = "binomial", type.measure = "auc", standardize = FALSE)
coefs <- coef(cv, s = "lambda.min")
print(coefs)
predictions <- predict(cv, newx = x, s = "lambda.min", type = "response")
class_predictions <- ifelse(predictions > 0.5, 1, 0)
accuracy <- mean(class_predictions == y)
sprintf("Accuracy: %f", accuracy)
i <- which(cv$lambda == cv$lambda.min)
sprintf("cv auc: %f", cv$cvm[i])
sprintf("sd: %f", cv$cvsd[i]/cv$cvm[1])
auc(y, predictions)
```

  Check AI alone
```{r}
set.seed(0)
norm_issues <- all_issues %>% filter(group %in% c("AI")) %>%mutate(across(where(is.numeric) & !merged, ~ as.vector(scale(.))))
norm_issues$group <- droplevels(norm_issues$group)

x <- model.matrix(merged ~ ., data = norm_issues %>% select(-vertices, -group))[,-1]
y <- (norm_issues)$merged
cv <- cv.glmnet(x, y, alpha = 0, family = "binomial", type.measure = "auc", standardize = FALSE)
coefs <- coef(cv, s = "lambda.min")
print(coefs)
predictions <- predict(cv, newx = x, s = "lambda.min", type = "response")
class_predictions <- ifelse(predictions > 0.5, 1, 0)
accuracy <- mean(class_predictions == y)
sprintf("Accuracy: %f", accuracy)
i <- which(cv$lambda == cv$lambda.min)
sprintf("cv auc: %f", cv$cvm[i])
sprintf("sd: %f", cv$cvsd[i]/cv$cvm[1])
auc(y, predictions)
```

  Do AI-Human regression
```{r}
set.seed(0)
all_issues$density[is.na(all_issues$density)] <- 0
norm_issues <- all_issues %>% filter(group %in% c("Human", "AI")) %>%mutate(across(where(is.numeric) & !merged, ~ as.vector(scale(.))))
norm_issues$group <- droplevels(norm_issues$group)

x <- model.matrix(merged ~ (. -group) * group, data = norm_issues %>% select(-vertices))[,-1]
y <- (norm_issues)$merged
cv <- cv.glmnet(x, y, alpha = 0, family = "binomial", type.measure = "auc", standardize = FALSE)
coefs <- coef(cv, s = "lambda.min")
print(coefs)

predictions <- predict(cv, newx = x, s = "lambda.min", type = "response")
class_predictions <- ifelse(predictions > 0.5, 1, 0)

accuracy <- mean(class_predictions == y)
sprintf("Accuracy: %f", accuracy)

i <- which(cv$lambda == cv$lambda.min)
sprintf("cv auc: %f", cv$cvm[i])
sprintf("sd: %f", cv$cvsd[i]/cv$cvm[1])

auc(y, predictions)
```
```{r}
coef_vec <- as.numeric(coefs)[-c(1,16)]
var_names <- (coefs@Dimnames[[1]])[2:15]

AI_coef <- coef_vec[1:14]
Human_coef <- AI_coef + coef_vec[15:28]

coef_data <-data.frame(row.names = var_names, AI = AI_coef, Human = Human_coef)
coef_data <- as.data.frame(t(coef_data))
```

```{r}
coef_plot <- ggplot(coef_data 
                    %>% rownames_to_column(var="group") 
                    %>% pivot_longer(-group),
                    aes(x=value, y=name)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey") +
  geom_line(aes(group=name)) +
  geom_point(aes(color=group), size = 3) +
  labs(x = "Estimated Effect Size on Merge Success",
       y = "Metric",
       color = "Agent") +
  theme_classic()
show(coef_plot)
ggsave(filename="./plots/pool_interaction_AI-Human.png", plot = coef_plot)
```
  Add Codex and Devin into the regression
```{r}
set.seed(0)
norm_issues <- all_issues %>% filter(group %in% c("Human", "AI", "Devin", "OpenAI Codex")) %>% mutate(across(where(is.numeric) & !merged, ~ as.vector(scale(.))))
norm_issues$group <- droplevels(norm_issues$group)

x <- model.matrix(merged ~ (. -group) * group, data = norm_issues %>% select(-vertices))[,-1]
y <- (norm_issues)$merged
cv <- cv.glmnet(x, y, alpha = 0, family = "binomial", type.measure = "auc", standardize = FALSE)
coefs <- coef(cv, s = "lambda.min")
print(coefs)

predictions <- predict(cv, newx = x, s = "lambda.min", type = "response")
class_predictions <- ifelse(predictions > 0.5, 1, 0)

accuracy <- mean(class_predictions == y)
sprintf("Accuracy: %f", accuracy)

i <- which(cv$lambda == cv$lambda.min)
sprintf("cv auc: %f", cv$cvm[i])
sprintf("sd: %f", cv$cvsd[i]/cv$cvm[1])

auc(y, predictions)
```

```{r}
coef_vec <- as.numeric(coefs)[-c(1,16,17,18)]
var_names <- (coefs@Dimnames[[1]])[2:15]

AI_coef <- coef_vec[1:14]
Devin_coef <- AI_coef + coef_vec[seq(from = 15, to = length(coef_vec), by = 3)]
Human_coef <- AI_coef + coef_vec[seq(from = 16, to = length(coef_vec), by = 3)]
Codex_ceof <- AI_coef + coef_vec[seq(from = 17, to = length(coef_vec), by = 3)]

coef_data <-data.frame(row.names = var_names, AI = AI_coef, Human = Human_coef, Devin = Devin_coef, Codex = Codex_ceof)
coef_data <- as.data.frame(t(coef_data))

```

```{r}

plot_data <- coef_data %>% 
  rownames_to_column(var = "group") %>% 
  pivot_longer(-group) %>%
  mutate(
    y_base = as.numeric(as.factor(name)),
    offset = case_when(
      group == "AI" ~ 0.25,
      group == "Human" ~ 0.25,
      group == "Devin" ~ -0.25,
      group == "Codex" ~ -0.25,
      TRUE ~ 0.0
    ),

    y_custom = y_base + offset
  )

coef_plot <- ggplot(plot_data, aes(x = value, y = y_custom)) +
  geom_hline(yintercept = seq(0.5, ncol(coef_data) + 0.5, by = 1), 
             color = "gray80", 
             linewidth = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey") +
  geom_line(data = filter(plot_data, group %in% c("AI", "Human")),
            aes(y = y_custom, group = name), 
            color = "gray20") +
  geom_point(aes(color = fct_relevel(group, "AI", "Codex" ,"Human")), size = 2.3) +
  scale_y_continuous(
    breaks = seq_along(unique(plot_data$name)),
    labels = levels(as.factor(plot_data$name))) +
  labs(x = "Estimated Effect Size on Merge Success",
       y = "Metric",
       color = "Agent") +
  theme_classic()
show(coef_plot)
ggsave(filename="./plots/agent_coef_plot.png", plot = coef_plot)
```